### 2019-11-26 paper.rmd
update

---
title             : "Getting started with Latent Profile Analysis (LPA) in the psychological sciences: A tutorial using the tidyLPA R package that provides an interface to open-source and commercial software"
shorttitle        : "Title"

author: 
  - name          : "Joshua Rosenberg"
    affiliation   : "1"
    corresponding : yes
    address       : "1122 Volunteer Blvd., Knoxville, TN, 37996"
    email         : "jmrosenberg@utk.edu"
  - name          : "Caspar van Lissa"
    affiliation   : "2"
  - name          : "Jennifer Schmidt"
    affiliation   : "3"
  - name          : "Patrick Beymer"
    affiliation   : "3"
  - name          : "Daniel Anderson"
    affiliation   : "4"
  - name          : "Matthew Schell"
    affiliation   : "3"
  - name          : "Rebecca Steingut"
    affiliation   : "5"
  - name          : "Stephanie Wormington"
    affiliation   : "6"

affiliation:
  - id            : "1"
    institution   : "University of Tennessee, Knoxville"
  - id            : "2"
    institution   : "Utrecht University"
  - id            : "3"
    institution   : "Michigan State University"
  - id            : "4"
    institution   : "University of Oregon"
  - id            : "5"
    institution   : "Columbia University"
  - id            : "6"
    institution   : "University of Virginia"

@@

The questions on the Scimo per and post survey are from Ryan and Deci's "Learning Climate Questionnaire".  You can read a short description here:

https://selfdeterminationtheory.org/pas-learning-climate/

It is 15 questions, but they identify 6 which can be used as  short-form.  Questions 1, 2, 4, 7, 10, 14.  I have attached the PDF of the survey since you have to create an account on their website do download it.  Below are the question text, along with the variable name in the Scimo data:

I feel that my instructor provides me choices and options.  t_choices1
I feel understood by my instructor.  t_understand1
My instructor conveyed confidence in my ability to do well in the course.  t_confidence1
My instructor encouraged me to ask questions.  t_encourage1
My instructor listens to how I would like to do things.  t_listen1
My instructor tries to understand how I see things before suggesting a new way to do things.  t_triesunder1

Also, is this tutorial going to be a journal paper, or something on github, etc. as a kind of manual for using LPA?  If so, and you think I would be a good fit, I would love to help out on it if you have author space!

Let me know if you need anything else.

@@

* `estimate_profiles()` now estimates any number of profiles, for any number of models, using any algorithm. It has a simplified user interface. This makes it a more general function, of which the former version (which could only estimate one number of profiles for one model) is a special, restricted case.

* `compare_solutions()` now compares solutions estimated by estimate_profiles(), suggesting the optimal model and number of classes. This replaces its dual role of estimating AND comparing solutions.
    `* `compare_solutions()` gives you the same set of fit indices for Mclust and Mplus. It provides a broader range of fit indices than before
    * `compare_solutions()` calls `AHP()``, a function which uses the Automatic Hierarchical Process to determine the optimal number of classes, based on several reputable fit indices

* `plot_profiles()` now incorporates best practices for plotting latent profile analyses: Visualizing cluster centroids, standard errors, cluster variances, and classification uncertainty by means of weighted raw data

* `impute_missing()` offers two easy-to-use single imputation options, for use with the Mclust package (which does not natively handle missing data). Mplus handles partially missing data by default.

@@

## Example analysis using estimate_profiles()

@@

```{r}
library(tidyverse)
library(tidyLPA)

pisaUSA15[1:100, ] %>%
    select(broad_interest, enjoyment, self_efficacy) %>%
    single_imputation() %>%
    estimate_profiles(3)
```

@@

```{r, eval = FALSE}
pisaUSA15[1:100, ] %>%
    select(broad_interest, enjoyment, self_efficacy) %>%
    single_imputation() %>%
    estimate_profiles(3, package = "MplusAutomation")
```

@@

```{r, eval = FALSE}
pisaUSA15[1:100, ] %>%
    select(broad_interest, enjoyment, self_efficacy) %>%
    single_imputation() %>%
    scale() %>%
    estimate_profiles(3) %>% 
    plot_profiles()
```

## Example analysis using compare_solutions()

@@

```{r, eval = TRUE}
pisaUSA15[1:100, ] %>%
    select(broad_interest, enjoyment, self_efficacy) %>%
    single_imputation() %>%
    estimate_profiles(1:3, 
                      variances = c("equal", "varying"),
                      covariances = c("zero", "varying")) %>%
    compare_solutions(statistics = c("AIC", "BIC"))
```

@@

* `compare_solutions()` no longer estimates profiles. It merely compares solutions estimated by `estimate_profiles()`

* all Mplus-specific functions are deprecated. `estimate_profiles()` offers a unified interface to Mclust and Mplus. These Mplus-specific functions are deprecated:

@@

* estimate_profiles and compare_solutions no longer select variables, passed as unquoted variable names, from 'df'. Instead, they use all variables in df. To select variables prior to analysis, use:

@@

* estimate_profiles and compare_solutions no longer center and scale the raw data. To center and scale variables prior to analysis, use scale(df, center = TRUE, scale = TRUE)

* estimate_profiles no longer accepts arguments specific to Mplus or Mclust. Instead, you can specify additional arguments directly to the core functions `mplusModeler()` and `Mclust()`. See the help file for these functions for more details. Deprecated arguments, whose functionality is supplanted by passing arguments directly to these core functions, are:

@@

# tidyLPA function with _mplus() appended require Mplus to use

```{r, eval = FALSE}
estimate_profiles_mplus(data, var1, var2, var3, ariances = “free”, covariances = “equal”) %>% 
    plot_profiles_mplus()
```

>>>>

---
title             : "Doing Latent Profile Analysis in a more rigorous and reproducible way: A tutorial using the tidyLPA R package"
shorttitle        : "Title"

author: 
  - name          : "Joshua Rosenberg"
    affiliation   : "1"
    corresponding : yes
    address       : "1122 Volunteer Blvd., Knoxville, TN, 37996"
    email         : "jmrosenberg@utk.edu"
  - name          : "Caspar van Lissa"
    affiliation   : "2"
  - name          : "Jennifer Schmidt"
    affiliation   : "3"
  - name          : "Patrick Beymer"
    affiliation   : "3"
  - name          : "Daniel Anderson"
    affiliation   : "4"
  - name          : "Matthew Schell"
    affiliation   : "3"

affiliation:
  - id            : "1"
    institution   : "University of Tennessee, Knoxville"
  - id            : "2"
    institution   : "Utrecht University"
  - id            : "3"
    institution   : "Michigan State University"
  - id            : "4"
    institution   : "University of Oregon"
  - id            : "5"

@@

<!-- The questions on the Scimo per and post survey are from Ryan and Deci's "Learning Climate Questionnaire".  You can read a short description here: -->

<!-- https://selfdeterminationtheory.org/pas-learning-climate/ -->

<!-- It is 15 questions, but they identify 6 which can be used as  short-form.  Questions 1, 2, 4, 7, 10, 14.  I have attached the PDF of the survey since you have to create an account on their website do download it.  Below are the question text, along with the variable name in the Scimo data: -->

<!-- I feel that my instructor provides me choices and options.  t_choices1 -->
<!-- I feel understood by my instructor.  t_understand1 -->
<!-- My instructor conveyed confidence in my ability to do well in the course.  t_confidence1 -->
<!-- My instructor encouraged me to ask questions.  t_encourage1 -->
<!-- My instructor listens to how I would like to do things.  t_listen1 -->
<!-- My instructor tries to understand how I see things before suggesting a new way to do things.  t_triesunder1 -->

@@

  - `estimate_profiles()` now estimates any number of profiles, for any number
    of models, using any algorithm. It has a simplified user interface. This
    makes it a more general function, of which the former version (which could
    only estimate one number of profiles for one model) is a special, restricted
    case.

  - `compare_solutions()` now compares solutions estimated by
    estimate\_profiles(), suggesting the optimal model and number of classes.
    This replaces its dual role of estimating AND comparing solutions. ` *
     `compare\_solutions()\` gives you the same set of fit indices for Mclust
    and Mplus. It provides a broader range of fit indices than before
      - `compare_solutions()` calls \`AHP()\`\`, a function which uses the
        Automatic Hierarchical Process to determine the optimal number of
        classes, based on several reputable fit indices

  - `plot_profiles()` now incorporates best practices for plotting latent
    profile analyses: Visualizing cluster centroids, standard errors, cluster
    variances, and classification uncertainty by means of weighted raw data

  - `impute_missing()` offers two easy-to-use single imputation options, for use
    with the Mclust package (which does not natively handle missing data). Mplus
    handles partially missing data by default.

@@

## Example analysis using estimate\_profiles()

@@

``` {r}
library(tidyverse)
library(tidyLPA)
```

pisaUSA15[1:100, ] %\>% select(broad\_interest, enjoyment, self\_efficacy)
%\>% single\_imputation() %\>% estimate\_profiles(3)

``` 
```

@@

``` {r, eval = FALSE}
pisaUSA15[1:100, ] %>%
    select(broad_interest, enjoyment, self_efficacy) %>%
    single_imputation() %>%
    estimate_profiles(3, package = "MplusAutomation")
```

@@

``` {r, eval = FALSE}
pisaUSA15[1:100, ] %>%
    select(broad_interest, enjoyment, self_efficacy) %>%
    single_imputation() %>%
    scale() %>%
    estimate_profiles(3) %>% 
    plot_profiles()
```

## Example analysis using compare\_solutions()

@@

``` {r, eval = TRUE}
pisaUSA15[1:100, ] %>%
    select(broad_interest, enjoyment, self_efficacy) %>%
    single_imputation() %>%
    estimate_profiles(1:3, 
                      variances = c("equal", "varying"),
                      covariances = c("zero", "varying")) %>%
    compare_solutions(statistics = c("AIC", "BIC"))
```

@@

  - `compare_solutions()` no longer estimates profiles. It merely compares
    solutions estimated by `estimate_profiles()`

  - all Mplus-specific functions are deprecated. `estimate_profiles()` offers a
    unified interface to Mclust and Mplus. These Mplus-specific functions are
    deprecated:

@@

  - estimate\_profiles and compare\_solutions no longer select variables, passed
    as unquoted variable names, from 'df'. Instead, they use all variables in
    df. To select variables prior to analysis, use:

@@

  - estimate\_profiles and compare\_solutions no longer center and scale the raw
    data. To center and scale variables prior to analysis, use scale(df, center
    = TRUE, scale = TRUE)

  - estimate\_profiles no longer accepts arguments specific to Mplus or Mclust.
    Instead, you can specify additional arguments directly to the core functions
    `mplusModeler()` and `Mclust()`. See the help file for these functions for
    more details. Deprecated arguments, whose functionality is supplanted by
    passing arguments directly to these core functions, are:

@@

# tidyLPA function with \_mplus() appended require Mplus to use

``` {r, eval = FALSE}
estimate_profiles_mplus(data, var1, var2, var3, ariances = “free”, covariances = “equal”) %>% 
    plot_profiles_mplus()
```
